Ants-Review: A Protocol For Open Anonymous Peer-Reviews

### 2020-09-21 paper-v1.tex
Abstract v 2.

Peer-review is a necessary and essential quality control step for scientific publications but lacks proper  incentives. Indeed, the process, which is very costly in terms of time and intellectual investment, not only is not remunerated by the journals but it is also not openly recognized by the academic community as a relevant scientific output for a researcher. Therefore, scientific dissemination is affected in the timeliness, quality and fairness. Here, to solve this issue, we propose a blockchain-based incentive system that rewards scientists for peer-reviewing other scientists’ work and that builds up trust and reputation. We designed a protocol of smart contracts called Ants-Review that allows authors to issue a bounty for open anonymous peer-reviews. If requirements are met, peer-reviews will be accepted and payed by the approver proportionally to their quality and value. To promote ethical behaviour and inclusiveness the system implements a gamified mechanism that allows the whole community to evaluate the peer-reviews, vote for the best ones and share the incentives with the referees.
\keywords{Blockchain \and Open Science \and Peer-review \and Incentivization.}

@@

\section{Background}
\subsection{Peer-review: present problems and mild solutions}
Peer review is still the only quality control mechanism devoted to evaluate scientific outcomes. The purpose of peer review is "improving the quality of the published paper, determining the originality of the manuscript, determining the importance of the findings, detecting fraud, and detecting plagiarism." \cite{Gropp-PeerRevStress}. However, the system is 'ﬂawed' and outdated \cite{Smith2006} and presents multifaceted issues, here reviewed.
\paragraph{A slow multi-stage process.} The main issues affecting the effectiveness of peer-review is the delay between paper submission and journal acceptance for publication. The traditional peer-review process is completely controlled by the editor(s). The author(s) submit manuscript to the journal where an editorial team assesses if the paper meets the scopes of the journal and novelty criteria. If the editorial decision is to send the manuscript for review, the handling editor personally selects potential reviewers. The reviewers identity is usually only known to the editor but is hidden to the authors or the other reviewers themselves (single-blind review). Reviewers independently conduct their reviews by exposing in their reports strengths and weaknesses of the manuscript and sometimes substantially improving the draft. Depending on if the decision is a major or minor revision, authors are invited to re-submit a corrected or improved version of the manuscript with a letter to the reviewers. The same reviewers might be contacted again to continue the same peer-review process. This process can take multiple rounds and is a huge time investment both for authors and reviewers. An analysis of all papers published in PubMed for a time period of 30 years claims that the median review time is around 100 days\cite{Kendall-peerrev}.
\paragraph{Lack of recognition.} Peer-reviewing is an invisible activity purely conducted on voluntary basis, neither paid by the journals or officially credited via standard scientific metrics (such as the ones that establish the 'impact factor' of an author). Thus, it does not lead to advancements in career or help securing grants. Researchers are motivated to do peer-review by a sense of belonging and a desire to 'give back' to the community. A major consequence of not promoting incentives for the quality (and quantity) of peer-reviews is to either slow down publication of potentially good research which awaits for validation or let bad science be published through sloppy and uncritical reviews.
\paragraph{Fraud and misconduct.} Due to the 'publish or perish culture' pressures, unethical behaviour from reviewers has been occasionally reported, from abusive behaviour towards authors \cite{Smith2006,tragedy-reviewers} to identity fraud. Some studies have reported an improvement in the transparency and civility of the review process when open reports are released according to the standards of open peer review.
\paragraph{Social and cognitive biases.} Given the fact that anonymity is usually asymmetrically applied only for reviewers, many power related dynamics can influence the reviewers decision, such as gender or cultural discrimination and social prestige of the institution. To solve this problem some journals have implemented double-blind review process (identity of both authors and reviewers are masked) which seems to reduce the bias towards minorities.
\paragraph{Peer reviews need to be...reviewed.} There is high variability in the reliability and depth of reviews and a recurrent question is: "Who watches the watchers?".
\paragraph{Need for more reviewers.} There is a disproportion between the progressive increase in journals publications and the number of experienced reviewers selected for the task which demands an expansion of the reviewer's pool including early career scientists. \cite{tragedy-reviewers}.

@@

\section{System concept}
In this paper we propose Ants-Review, a new incentivisation mechanism built on a distributed platform that issues open peer-reviews to validate scientific papers while preserving the anonymity of its contributors. We imagine a final paper originating from the peer-review process as a complex system that emerges from the interactions between the authors and the reviewers, a whole that is more than the sum of its parts. Therefore, the name evokes an ant colony as a self-organising organism in which all micro-contributions of the individuals emerge into complex behaviour.
\newline The original proposal behind this paper can be found here \cite{AntsReview}.
\newline Its design and implementation are exposed in the following section.

>>>>

Peer-review is a necessary and essential quality control step for scientific publications but lacks proper  incentives. Indeed, the process, which is very costly in terms of time and intellectual investment, not only is not remunerated by the journals but it is also not openly recognized by the academic community as a relevant scientific output for a researcher. Therefore, scientific dissemination is affected in the timeliness, quality and fairness. Here, to solve this issue, we propose a blockchain-based incentive system that rewards scientists for peer-reviewing other scientists’ work and that builds up trust and reputation. We designed a protocol of smart contracts called Ants-Review that allows authors to issue a bounty for open anonymous peer-reviews on Ethereum. If requirements are met, peer-reviews will be accepted and payed by the approver proportionally to their assessed quality. To promote ethical behaviour and inclusiveness the system implements a gamified mechanism that allows the whole community to evaluate the peer-reviews and vote for the best ones.
\keywords{Blockchain \and Ethereum \and Peer-review \and Incentivization.}

@@

\section{Background}
\subsection{Peer-review: present problems and mild solutions}
Peer review is still the only quality control mechanism devoted to evaluate scientific outcomes. The purpose of peer review is "improving the quality of the published paper, determining the originality of the manuscript, determining the importance of the findings, detecting fraud, and detecting plagiarism." \cite{Gropp-PeerRevStress}. However, the system is 'ﬂawed' and outdated \cite{Smith2006} and presents multifaceted issues, here reviewed.
\paragraph{A slow multi-stage process.} The main issues affecting the effectiveness of peer-review is the delay between paper submission and journal acceptance for publication. The traditional peer-review process is completely controlled by the editor(s). The author(s) submits manuscript to the journal where an editorial team assesses if the paper meets the scopes of the journal and novelty criteria. If the editorial decision is to send the manuscript for review, the handling editor personally selects potential reviewers. The reviewers identity is usually only known to the editor but is hidden to the authors or the other reviewers themselves (single-blind review). Reviewers independently conduct their reviews by exposing in their reports strengths and weaknesses of the manuscript and sometimes substantially improving the draft. Depending on if the decision is a major or minor revision, authors are invited to re-submit a corrected or improved version of the manuscript. The same reviewers might be contacted again to continue the same peer-review process. This process can take multiple rounds and is a huge time investment both for authors and reviewers. An analysis of all papers published in PubMed for a time period of 30 years claims that the median review time is around 100 days\cite{Kendall-peerrev}.
\paragraph{Lack of recognition.} Peer-reviewing is an invisible activity purely conducted on voluntary basis, neither paid by the journals or officially credited via standard scientific metrics (such as the ones that establish the 'impact factor' of an author). Thus, it does not lead to advancements in career or help securing grants. Researchers are motivated to do peer-review by a sense of belonging and a desire to 'give back' to the community. A major consequence of not promoting incentives for the quality (and quantity) of peer-reviews is to either slow down publication of potentially good research which awaits for validation or let bad science be published through sloppy and uncritical reviews.
\paragraph{Fraud and misconduct.} Due to the 'publish or perish culture' pressures, unethical behaviour from reviewers has been occasionally reported, from abusive behaviour towards authors \cite{Smith2006,tragedy-reviewers} to identity fraud. Some studies have reported an improvement in the transparency and civility of the review process when open reports are released according to the standards of \emph{open peer review}.
\paragraph{Social and cognitive biases.} Given the fact that anonymity is usually asymmetrically applied only for reviewers, many power related dynamics can influence the reviewers decision, such as gender or cultural discrimination and social prestige of the institution. To solve this problem some journals have implemented \emph{double-blind review process} (identity of both authors and reviewers are masked) which seems to reduce the bias towards minorities.
\paragraph{Peer reviews need to be...reviewed.} There is high variability in the reliability and depth of reviews and a recurrent question is: "Who watches the watchers?".
\paragraph{Need for more reviewers.} There is a disproportion between the progressive increase in journals publications and the number of experienced reviewers selected for the task which demands an expansion of the reviewer's pool including early career scientists. \cite{tragedy-reviewers}.

@@

\section{System concept}
In this paper we propose Ants-Review, a new incentivisation mechanism built on Ethereum that issues open peer-reviews to validate scientific papers while preserving the anonymity of its contributors. We imagine a final paper originating from the peer-review process as a complex system that emerges from the interactions between the authors and the reviewers, a whole that is more than the sum of its parts. Therefore, the name evokes an ant colony as a self-organising organism in which all micro-contributions of the individuals emerge into complex behaviour.
\newline The original proposal behind this paper can be found here \cite{AntsReview}.
\newline Its design and implementation are exposed in the following section.
